Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-04-19 paper/ijcai16.tex
paper: squeeze equations a bit.

A straightforward encoding of the above desiderata leads to the
following mixed integer {\em non-linear} optimization problem over the
non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge \mu \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
%    \\
%        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
Let us illustrate the above piece by piece. The objective is composed
of four parts: we maximize the shared margin $\mu$ (first part) and
minimize the total sum of the ranking errors $\veps^i$ incurred by
each weight vector $\vw^{i}$ (second part), while at the same time
regularizing the magnitude of the weights (third part) and the quality
of the configurations $\{ \vx^{i} \}$ (last part). The non-negative
hyperparameters $\alpha,\beta,\gamma$ control the influence of the
various components. The weight regularization term copes with the
common scenario in which the user has strong preferences about some
attributes, but is indifferent to most of
them. %~\cite{} \stefano{@Paolo: add refs}.
The $\ell_1$ penalty is frequently used to improve the sparsity of
learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent
gains in generalization ability and efficiency, as confirmed by our
empirical findings (see
Section~\ref{sec:experiments}). Constraint~(\ref{eq:wyconstr})
enforces the correct ranking of the observed user preferences, while
%constraint
~(\ref{eq:wxconstr}) ensures that the generated
configurations are diverse in terms of the weight vectors they
maximize. Constraints~(\ref{eq:wbounds}) and~(\ref{eq:xbounds}) ensure
that the weights and configurations are feasible, while the last one
guarantees the non-negativity of the slacks.  Since we require
$\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the
weights to be non-negative.

@@

Substituting the above MILP constraints into the original non-linear
formulation, we obtain the following mixed-integer linear problem:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu
    \\
    %     & \;\; \forall \; i \in [k], \forall \; z \in [m] \nonumber
    % \\
        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [m] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \} \label{eq:volatile1}
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} \label{eq:volatile2}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \quad \veps^{i}_h \ge 0 \nonumber
%    \\
%        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
which can be solved by any suitable MILP solver.

>>>>

A straightforward encoding of the above desiderata leads to the
following mixed integer {\em non-linear} optimization problem over the
non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge \mu \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \;\;, \veps^{i} \ge 0 \label{eq:xbounds}
%    \\
%        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
%    \\
%        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
%    \\
%        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
Let us illustrate the above piece by piece. The objective is composed
of four parts: we maximize the shared margin $\mu$ (first part) and
minimize the total sum of the ranking errors $\veps^i$ incurred by
each weight vector $\vw^{i}$ (second part), while at the same time
regularizing the magnitude of the weights (third part) and the quality
of the configurations $\{ \vx^{i} \}$ (last part). The non-negative
hyperparameters $\alpha,\beta,\gamma$ control the influence of the
various components. The weight regularization term copes with the
common scenario in which the user has strong preferences about some
attributes, but is indifferent to most of
them. %~\cite{} \stefano{@Paolo: add refs}.
The $\ell_1$ penalty is frequently used to improve the sparsity of
learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent
gains in generalization ability and efficiency, as confirmed by our
empirical findings (see
Section~\ref{sec:experiments}). Constraint~(\ref{eq:wyconstr})
enforces the correct ranking of the observed user preferences, while
%constraint
~(\ref{eq:wxconstr}) ensures that the generated
configurations are diverse in terms of the weight vectors they
maximize. Constraints~(\ref{eq:wbounds}) and~(\ref{eq:xbounds}) ensure
that the weights and configurations are feasible
%, while the last one
and
guarantees the non-negativity of the slacks.  Since we require
$\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the
weights to be non-negative.

@@

Substituting the above MILP constraints into the original non-linear
formulation, we obtain the following mixed-integer linear problem:
%
{\footnotesize
\begin{align}
    \max
        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \quad \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu
    \\
    %     & \;\; \forall \; i \in [k], \forall \; z \in [m] \nonumber
    % \\
        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [m] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \} \label{eq:volatile1}
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} \label{eq:volatile2}
    \\
        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
    \\
        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \;, \veps^{i} \ge 0 \nonumber
%    \\
%        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
%    \\
%        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
%    \\
%        & \;\; \mu \ge 0 \nonumber
\end{align}
}
%
which can be solved by any suitable MILP solver.
